---
title: "Data_manipulation"
author: "Ada"
date: "9/20/2018"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
```

## Import FAS csv files

```{r}
options(tibble.print_min = 3)

litters_data = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```

specifying the columns you want to keep by naming all of them 

```{r}
select(litters_data, group, litter_number, gd0_weight, pups_born_alive)
select(litters_data, group:gd_of_birth)    # specify the specify a range of columns to keep

select(litters_data, -pups_survive)  #You can also specify columns you’d like to remove



```



## filter

```{r}
filter(litters_data, gd0_weight < 25)

filter(litters_data, pups_born_alive == 8)

filter(litters_data, ! is.na(gd0_weight))


filter(litters_data, group == "low8" | group == "low7")
filter(litters_data, group %in% c("low8", "low7"))

filter(litters_data, pups_born_alive %in% 4:6)

filter(litters_data, pups_born_alive %in% 4:6, !is.na(gd0_weight))


filter(pups_data, sex == 1)

```
"==" whether or not the two things are equal


## mutate

```{r}
mutate(litters_data,
  wt_gain = gd18_weight - gd0_weight,   #creating a new variable
  group = tolower(group)  # lower letter
)

mutate(litters_data,           
  wt_gain = gd18_weight - gd0_weight,   
  wt_gain_squared = wt_gain^2
)


```


## arrange

```{r}
arrange(litters_data, gd0_weight)
arrange(litters_data, desc(pups_born_alive), gd0_weight)
```

## %in%

```{r}

#1
litters_data_raw = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
litters_data_with_vars = mutate(litters_data_selected_cols, 
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group))
litters_data_with_vars

#2
litters_data_clean = 
  mutate(
    select(
      janitor::clean_names(
        read_csv("./data/FAS_litters.csv", col_types = "ccddiiii")
        ), 
    -pups_survive
    ),
  wt_gain = gd18_weight - gd0_weight,
  group = tolower(group)
  )
litters_data_clean

#3 using pipe   much easier to read
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = tolower(group))

litters_data

#the result of the 3 options are the same


```

Illustrate placeholder use with "lm"

```{r}
?lm

litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
   # wt_gain = gd18_weight - gd0_weight,
    group = tolower(group)) %in%
#    lm(gd18_weight ~ gd0_weight)
litters_data


#  what . mean?

