---
title: "Data_import"
author: "Ada"
date: "9/18/2018"
output: html_document
---
```{R setup, include = FALSE}
library(tidyverse)
```
## Import FAS csv files
```{r}
litters_data = read_csv(file = "~/Documents/Ada's Documents/Columbia Biostatistics/Data Science/R programming/Lecture_5/Data_wrangling_1/data/FAS_litters.csv")
litters_data = read_csv(file = "./data/FAS_litters.csv")
litters_data = read_csv("./data/FAS_litters.csv")
litters_data = janitor::clean_names(litters_data)
```

* ***the first two lines of code, are they the same?***
* janitor::clean_names:Resulting names are unique and consist only of the _ character, numbers, and letters. Doing so will take whatever the column names are and convert them to lower snake case.

##here::here
```{r}

library(here)
here("data", "file_i_want.csv")
```

## learning assessment 1
```{r learning assessment1}
pups_data = read_csv((file = "./data/FAS_pups.csv"))
pups_data = read_csv((file = "~/Documents/Ada's Documents/Columbia Biostatistics/Data Science/R programming/Lecture_5/Data_wrangling_1/data/FAS_pups.csv"))
```

```{r}
head(litters_data)
tail(litters_data)
skimr::skim(litters_data)
litters_data = read_csv(file = "./data/FAS_litters.csv",
  skip = 10, col_names = FALSE)

```

## Parsing columns
```{r}
litters_data = read_csv(file = "./data/FAS_litters.csv",
  col_types = cols(
    Group = col_character(),
    `Litter Number` = col_character(),
    `GD0 weight` = col_double(),
    `GD12 weight` = col_double(),
    `GD of Birth` = col_integer(),
    `Pups born alive` = col_integer(),
    `Pups dead @ birth` = col_integer(),
    `Pups survive` = col_integer()
  )
)

# in shorthand
litters_data = read_csv(file = "./data/FAS_litters.csv",
  col_types = "ccddiiii"
)
```
 
 ##learning assessment 2
```{r}


pups_data = read_csv("./data/FAS_pups.csv", col_types = "ciiiii")
skimr::skim(pups_data)


pups_base = read.csv("./data/FAS_pups.csv")
pups_readr = read_csv("./data/FAS_pups.csv")

View(pups_base)
View(pups_readr)

pups_base
pups_readr

pups_base$S
pups_readr$S

pups_base$Sex
pups_readr$Sex
```

* In short, read_csv produces tibbles which are very similar to the base R data frames produced by read.csv. However, tibbles have some features that can help prevent mistakes and unwanted behavior.

